Weather Data Integration and Assimilation System\\

### 2020-02-09 lit/sec_1-fews.tex
Before Test plan
Paper - update abstract

\section{The Delft-FEWS flow forecasting system}
 \acrshort{fews} is developed by Deltares in the Nethaland with targeted for use by operational forecasting agencies. The software is freely available to download and setup. Tutorials and User setup guide \cite{DeltaresPublicWikiDelft-FEWSGuide} are available with well documentations. For non-commercial users need to have the licence agreement to use the system. And get new system updates. However, investments in new developments in the system are normally paid for by its users. The code base of \acrshort{fews} is currently not fully open source. But configuration guides and generic adapter module integration code examples are available for the users.

Most if the models are using model-centric approach such as inputs need to be in the format of specific to the model. Also the outputs produced by the model are specific to the model and hard to utilize by another systems (e.g. FLO2D). For the \acrshort{fews} is using much more modular approach, hence it is easy to integrate models into the \acrshort{fews} system. Thus \acrshort{fews} can be consider as an integration framework or a middle ware for the models.

@@

The key elements of a forecasting system are summarized by Haggett \cite{Haggett1998AnWales} as four main steps;
\begin{enumerate}
    \item Detection
    \item Forecasting
    \item Dissemination and Warning
    \item Response
\end{enumerate}
Within these four steps, \acrshort{fews} focuses on the forecasting step. The primary objective of this step is to provide additional lead time through predictions of short term future hydro-meteorological conditions \cite{Werner2005FloodCatchments}. It is true that, with providing accurate predictions with greater lead time can reduce the level of destruction. In order to do the forecasting, the system should be capable of integrating real-time data from hydrological and meteorological observation networks, and the dissemination of prediction results through appropriate products to the warning process.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.7\textwidth]{fews/Schematic-structure-of-a-fl-ood-forecasting-system-showing-the-position-of-Delft-FEWS_W640.jpg}
    \caption{Schematic structure of a flood forecasting system, showing the position of \acrshort{fews} and links to other primary systems within the operational environment. \cite{Werner2013TheSystem} }
    \label{fi:fews_schematic}
\end{figure}
The foundation of \acrshort{fews} is data-centric, with a common data-model through which all components interact. All time series data (both scalar and gridded) are stored in this common data-model in a database. Modeling capabilities are then linked to the system through one of the interfaces provided to the data-model \cite{Werner2013TheSystem}. As mention in the paper, having a common data-model make easy to store data in an efficient way. Operations like reporting and sharing the data also make it easy for integrate models. But the problem with this approach is, handling the multiple data formats. \acrshort{fews} overcome this issue with introducing adapters for most of commonly used data formats. This is a good feature which enable users to easily import data into the system. But at the same time, it's adding additional complexity into the system.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{fews/Architecture-of-Delft-FEWS-showing-the-data-base-the-data-access-layers-and-examples-of_W640.jpg}\\
    \caption{Architecture of \acrshort{fews} showing the data base, the data access layers and examples of functional modules that communicate through the data access layer. \cite{Werner2013TheSystem} }
    \label{fi:fews_data_layer}
\end{figure}
Most of the library of functions provided work equally on scalar and gridded time series. As with all other modules this communicates with the database solely through the data access layer as shown in \ref{fi:fews_data_layer} \cite{Werner2013TheSystem}. As it shown in the figure, it is clearly showing that, the system is depends on single database and via the data access layer all the requests are coming to the database though a connection. It is possible to setup and connect to a enterprise level database with clustering and shading as a paid solution, in order to serve many requests as possible. But the design it self inherently suffer with the database bottleneck which effected on running a large set of timeseries or models.

@@

Exchanging data with the model is primarily through XML files. In some cases these XML files may become very large, which may lead to I/O bottlenecks and subsequent performance issues \cite{Werner2013TheSystem}. This issue is thoroughly discussed previous paragraphs and the authors of the \cite{Werner2013TheSystem} seems to be notice and accept it here. For overcome this issue they introduced the file-based exchange of data. This includes the use of binary-XML files, streaming files through memory, and the use of \acrshort{NetCDF}-CF files. But far as for users it seems to be adding more complexity and context in order to get higher performance via the system.

If particular adapter for a model is not present in the list of \acrshort{fews} adapter, user has the flexibility to implement an adapter for it. The effort of developing an adapter for a model code not previously integrated with \acrshort{fews} will vary depending on the complexity of the model I/O formats \cite{Werner2013TheSystem}. If user wants to use existing adapter and need to configure something not supported via the existing adapter, it is hard to get done. Also there are some cases it is hard to develop an adapter or running alongside with the \acrshort{fews}. As an example FLO2D model is using for hydrologic modeling and the model is only support only on Windows based operating systems. If you setup the system on Linux base operating system and using Linux based models, it is hard for users to come up with a solution for integrating the FLO2D models.
The final step and one of important feature is exporting the data from the system in useful manner. In the \acrshort{fews}, it is supporting three main forms of product generation.
\begin{itemize}
\item Generate web reports with graphs, tables as well as summary reports
\item Export time series in a variety of formats
\item External applications actively retrieve data from the system
\end{itemize}
The forecasting process is often a sequence of steps, starting with the import of data, a number of data processing and modeling steps, and culminating in the generation of products to be disseminated to the warning process. This is supported via the work flow process in the \acrshort{fews}. The scope of the \acrshort{wdias} is not focus on adding such a feature in the system, rather user have to use scientific work flow management system or come up with own version of work flow.

When running sequential steps within a work flow in \acrshort{fews}, each step is run for the full time window required (e.g. the lead time of the forecast) before moving on to the next step. This precludes tightly coupling of models, even explicitly \cite{Werner2013TheSystem}. As it mentioned above, users can run multiple models parallel or run those sequentially as needed.

>>>>

\section{Delft-FEWS flow forecasting system}
\acrshort{fews} is developed by Deltares in the Nethaland with targeted for use by operational forecasting agencies. %The software is freely available to download and setup. Tutorials and User setup guide \cite{DeltaresPublicWikiDelft-FEWSGuide} are available with well documentations. For non-commercial users need to have the licence agreement to use the system. And get new system updates. However, investments in new developments in the system are normally paid for by its users. 
The code base of \acrshort{fews} is currently not fully open source. %But configuration guides and generic adapter module integration code examples are available for the users.
Most if the models are using model-centric approach such as inputs need to be in the format of specific to the model. Also, the outputs produced by the model are specific to the model and hard to utilize by another systems (e.g., FLO2D). As \acrshort{fews} %is using much more 
uses a modular approach it is easier to integrate new models.% into the \acrshort{fews} system. 
Thus, \acrshort{fews} can be consider as an integration framework or a middleware for the models.
\dbc{Cite Fews. Provide only the key details. It's sufficient say whether its commercial or open source. Do this for all other solutions discussed. Middleware is a single word. Check other places.}

@@

Are proposed by Haggett \cite{Haggett1998AnWales} key elements of a forecasting system include detection, forecasting, dissemination and warning, and response. Within these four steps, \acrshort{fews} focuses on the forecasting step. The primary objective of this step is to provide additional lead time through predictions of short term future hydro-meteorological conditions \cite{Werner2005FloodCatchments}. It is true that, with providing accurate predictions with greater lead time can reduce the level of destruction. In order to do the forecasting, the system should be capable of integrating real-time data from hydrological and meteorological observation networks, and the dissemination of prediction results through appropriate products to the warning process.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.7\textwidth]{fews/Schematic-structure-of-a-fl-ood-forecasting-system-showing-the-position-of-Delft-FEWS_W640.jpg}
    \caption{Schematic structure of a flood forecasting system including \acrshort{fews} and links to other primary systems within the operational environment \cite{Werner2013TheSystem}.}
    \label{fi:fews_schematic}
\end{figure}
\dbc{Most figures are of low quality. You need to make sure figure resolution/quality is not lost. Dot should come after reference no. Fix all places.}

The foundation of \acrshort{fews} is data-centric, with a common data-model through which all components interact. All time series data (both scalar and gridded) are stored in this common data-model in a database. Modeling capabilities are then linked to the system through one of the interfaces provided to the data-model \cite{Werner2013TheSystem}. As mention in the paper, having a common data-model make easy to store data in an efficient way. Operations like reporting and sharing the data also make it easy for integrate models. But the problem with this approach is, handling the multiple data formats. \acrshort{fews} overcome this issue with introducing adapters for most of commonly used data formats. This is a good feature which enable users to easily import data into the system. But at the same time, it's adding additional complexity into the system.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{fews/Architecture-of-Delft-FEWS-showing-the-data-base-the-data-access-layers-and-examples-of_W640.jpg}\\
    \caption{Architecture of \acrshort{fews} % showing the data base, the data access layers and examples of functional modules that communicate through the data access layer. 
    \cite{Werner2013TheSystem} }
    \label{fi:fews_data_layer}
\end{figure}
Most of the library of functions provided work equally on scalar and gridded time series. As with all other modules this communicates with the database solely through the data access layer as shown in \ref{fi:fews_data_layer} \cite{Werner2013TheSystem}. As it shown in the figure, it is clearly showing that, the system is depends on single database and via the data access layer all the requests are coming to the database though a connection. It is possible to setup and connect to a enterprise level database with clustering and shading as a paid solution, in order to serve many requests as possible. But the design it self inherently suffer with the database bottleneck which effected on running a large set of timeseries or models.

@@

Exchanging data with the model is primarily through XML files. In some cases these XML files may become very large, which may lead to I/O bottlenecks and subsequent performance issues \cite{Werner2013TheSystem}. This issue is thoroughly discussed previous paragraphs and the authors of the \cite{Werner2013TheSystem} seems to be notice and accept it here. For overcome this issue they introduced the file-based exchange of data. This includes the use of binary-XML files, streaming files through memory, and the use of \acrshort{netCDF} files. But far as for users it seems to be adding more complexity and context in order to get higher performance via the system.

If particular adapter for a model is not present in the list of \acrshort{fews} adapter, user has the flexibility to implement an adapter for it. The effort of developing an adapter for a model code not previously integrated with \acrshort{fews} will vary depending on the complexity of the model I/O formats \cite{Werner2013TheSystem}. If user wants to use existing adapter and need to configure something not supported via the existing adapter, it is hard to get done. Also there are some cases it is hard to develop an adapter or running alongside with the \acrshort{fews}. As an example FLO2D model is using for hydrologic modeling and the model is only support only on Windows based operating systems. If you setup the system on Linux base operating system and using Linux based models, it is hard for users to come up with a solution for integrating the FLO2D models.
The final step and one of important feature is exporting the data from the system in useful manner. %\acrshort{fews} supports three main forms of product generation, namely:
%\begin{itemize}
%\item Generate web reports with graphs, tables as well as summary reports
%\item Export time series in a variety of formats
%\item External applications actively retrieve data from the system
%\end{itemize}
The forecasting process is often a sequence of steps, starting with the import of data, a number of data processing and modeling steps, and culminating in the generation of products to be disseminated to the warning process. This is supported via the work flow process in the \acrshort{fews}. The scope of the \acrshort{wdias} is not focus on adding such a feature in the system, rather user have to use scientific work flow management system or come up with own version of work flow.

When running sequential steps within a work flow in \acrshort{fews}, each step is run for the full time window required (e.g. the lead time of the forecast) before moving on to the next step. This precludes tightly coupling of models, even explicitly \cite{Werner2013TheSystem}. As it mentioned above, users can run multiple models parallel or run those sequentially as needed. 
\dbc{It's "e.g.,". Fix all places. }

### 2020-02-09 lit/sec_2-lead.tex
Added All 1440 result graphs
Paper - update abstract

The reason behind developing \acrfull{lead} by US National Science Foundation (NSF)  is to addresses the fundamental research challenges needed to create an integrated, scalable framework for adaptive analyzing and predicting the atmosphere \cite{Droegemeier2005Service-OrientedWeather}. In order to predict and analysis weather models by researchers, it required many resources. Rather than each researcher is running and handling own computer resources to do the weather experiments, \acrshort{lead} is providing pool of resources, then the researchers can use this resource pool to run their experiment in shorter amounts of time and in higher scale. At the time, researchers are developing their experiment flow, they are not using the resources many, and other are using it at the same time. LEAD’s foundation is dynamic work flow orchestration and data management in a Web services framework \cite{Droegemeier2005Service-OrientedWeather}.

The most relevant and important analysis for designing \acrshort{lead} is understanding and analyzing the \acrfull{lead} architecture of it. LEAD’s complex array of services, applications, interfaces, and local and remote computing, networking, and storage resources is assembled by users in work flows to study mesoscale weather as it evolves \cite{Droegemeier2005Service-OrientedWeather}. As it follows the \acrshort{soa}, everything is implemented as independent services. This enable \acrshort{lead} to scale each service as required, and update each service without affecting other services. New models which are required for implementing a new forecast flow also has to be implement as a service and then integrate into the system.

In the \ref{fi:lead_system} shows the fundamental capabilities of the \acrshort{lead} system. 
From the high-level view, \acrshort{lead} lets users query for and acquire information, simulate and predict weather by using numerical atmospheric models, assimilate data, analyze, mine, and visualize data and model output.

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lead/LEAD-system-Fundamental-capabilities-familiar-to-meteorologists-are-shown-in-the-top_W640.jpg}\\
    \caption{LEAD system. Fundamental capabilities familiar to meteorologists are shown in the top level, below which are the associated tools for enacting these capabilities and the middle-ware that links everything together. System-generated products appear at the bottom. \cite{Droegemeier2005Service-OrientedWeather} }
    \label{fi:lead_system}
\end{figure}

The second level in \ref{fi:lead_system} contains the fundamental tools that help to link services together. This include following tools in brief \cite{Droegemeier2005Service-OrientedWeather};
\begin{itemize}
    \item a Web portal, the primary (though not exclusive) user entry point
    \item the ARPS Data Assimilation System (ADAS), a sophisticated tool for data quality control and assimilation
    \item myLEAD, a flexible metadata catalog service
    \item \acrfull{wrf} a next-generation atmospheric prediction and simulation model
    \item ADaM (Algorithm Development and Mining), a powerful suite of tools for mining observational data, assimilated data sets, and model output 
    \item Integrated Data Viewer (IDV) 
\end{itemize}

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lead/LEAD-system-framework-LEAD-is-composed-of-several-interacting-subsystems-with-the-LEAD_W640.jpg}\\
    \caption{ LEAD system framework. LEAD is composed of several interacting subsystems, with the LEAD grid representing a stable, secure environment for development and testing. \cite{Droegemeier2005Service-OrientedWeather} }
    \label{fi:lead_framework}
\end{figure}

@@

Lead System is consist of following sub components, as shown in the \ref{fi:lead_soa}. As a brief summary of the subsystems is \cite{Droegemeier2005Service-OrientedWeather};
\begin{itemize}
\item User subsystem - comprises the LEAD portal and enable user can access services
\item Data subsystem - handles data and metadata, any numerical model output produced by operational or experimental models, and user generated information.
\item Tools subsystem - consists of all meteorological and IT tools
\item Orchestration subsystem - provides the technologies that let users manage data flows and model execution streams, and create and mine output. It also provides linkages to other software and processes for continuous or on-demand applications. 
\end{itemize}

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.9\textwidth]{lead/LEADs-service-oriented-architecture-A-wide-variety-of-services-and-resources-grouped_W640.jpg}\\
    \caption{LEAD’s service-oriented architecture. A wide variety of services and resources, grouped here according to general functionality, comprise the LEAD architecture, including services that cut across all aspects of the system. \cite{Droegemeier2005Service-OrientedWeather} }
    \label{fi:lead_soa}
\end{figure}

The distributed computing systems in the LEAD grid represent a distributed test bed for developing, integrating, and testing LEAD’s components.

>>>>

%The reason behind developing \acrfull{lead} by US National Science Foundation (NSF)  is to
\acrfull{lead} addresses the fundamental research challenges needed to create an integrated, scalable framework for adaptive analyzing and predicting the atmosphere \cite{Droegemeier2005Service-OrientedWeather}. To predict and analysis weather models by researchers, it required many resources. Rather than each researcher is running and handling own computer resources to do the weather experiments, \acrshort{lead} is providing pool of resources, then the researchers can use this resource pool to run their experiment in shorter amounts of time and in higher scale. At the time, researchers are developing their experiment flow, they are not using the resources many, and other are using it at the same time. LEAD’s foundation is dynamic work flow orchestration and data management in a Web services framework \cite{Droegemeier2005Service-OrientedWeather}.

%The most relevant and important analysis for designing \acrshort{lead} is understanding and analyzing the \acrfull{lead} architecture of it. 
LEAD's complex array of services, applications, interfaces, and local and remote computing, networking, and storage resources is assembled by users in work flows to study mesoscale weather as it evolves \cite{Droegemeier2005Service-OrientedWeather}. As it follows the \acrshort{soa}, everything is implemented as independent services. This enable \acrshort{lead} to scale each service as required, and update each service without affecting other services. New models which are required for implementing a new forecast flow also has to be implement as a service and then integrate into the system.
\dbc{Check the use of apostrophe s. You need to use ' not ’.}

Figure \ref{fi:lead_system} shows the fundamental capabilities of the \acrshort{lead} system. From the high-level view, \acrshort{lead} lets users query for and acquire information, simulate and predict weather by using numerical atmospheric models, assimilate data, analyze, mine, and visualize data and model output.
\dbc{Make sure each reference to a figure or table is given as Figure x.y and Table x.y.}

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lead/LEAD-system-Fundamental-capabilities-familiar-to-meteorologists-are-shown-in-the-top_W640.jpg}\\
    \caption{Layered architecture of LEAD. Fundamental capabilities familiar to meteorologists are shown in the top level, next are the tools for enacting these capabilities and the middleware that links everything together. System-generated products appear at the bottom \cite{Droegemeier2005Service-OrientedWeather}. }
    \label{fi:lead_system}
\end{figure}

\dbc{Fig. 2.4 should use full text width.}

The second level contains the fundamental tools that help to link services together. This include following tools \cite{Droegemeier2005Service-OrientedWeather};
\begin{itemize}
    \item a Web portal, the primary (though not exclusive) user entry point
    \item the ARPS Data Assimilation System (ADAS), a sophisticated tool for data quality control and assimilation
    \item myLEAD, a flexible metadata catalog service
    \item \acrfull{wrf} a next-generation atmospheric prediction and simulation model
    \item ADaM (Algorithm Development and Mining), a powerful suite of tools for mining observational data, assimilated data sets, and model output 
    \item Integrated Data Viewer (IDV) 
\end{itemize}

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lead/LEAD-system-framework-LEAD-is-composed-of-several-interacting-subsystems-with-the-LEAD_W640.jpg}\\
    \caption{LEAD system framework %. LEAD is composed of several interacting subsystems, with the LEAD grid representing a stable, secure environment for development and testing. 
    \cite{Droegemeier2005Service-OrientedWeather} .}
    \label{fi:lead_framework}
\end{figure}

@@

As shown in the \ref{fi:lead_soa} LEAD consists of the following sub components:
\begin{itemize}
\item User subsystem -- comprises the LEAD portal and enable user can access services
\item Data subsystem -- handles data and metadata, any numerical model output produced by operational or experimental models, and user generated information.
\item Tools subsystem -- consists of all meteorological and IT tools
\item Orchestration subsystem -- provides the technologies that let users manage data flows and model execution streams, and create and mine output. It also provides linkages to other software and processes for continuous or on-demand applications. 
\end{itemize}

\begin{figure}[htp]
    \centering
    \includegraphics[width=1\textwidth]{lead/LEADs-service-oriented-architecture-A-wide-variety-of-services-and-resources-grouped_W640.jpg}\\
    \caption{LEAD’s service-oriented architecture %. A wide variety of services and resources, grouped here according to general functionality, comprise the LEAD architecture, including services that cut across all aspects of the system. 
    \cite{Droegemeier2005Service-OrientedWeather}.}
    \label{fi:lead_soa}
\end{figure}

The distributed computing systems in the LEAD grid represent a distributed test bed for developing, integrating, and testing LEAD’s components.
\dbc{Try to avoid singe sentence paras.}

### 2020-02-09 method/sec_1-soa.tex
Paper - update abstract

\section{Service Oriented Architecture}

In the initial phase of WDIAS architecture, it proposed with a \acrfull{soa} as shown in the \ref{fi:proposed_soa}. Mainly it consist of 4 modules;
- Import modules
Import modules are using for integration of data from different source which come in different data formats.
- Export modules
Export modules are using for data dissemination with different format which may required for different type of operations.
- Extension modules
Extension modules are capable of providing the data assimilation functionalities by providing some of key functionalities such as transformation, interpolation and validation. While data assimilation, existing timeseries may be modified or create new timeseries based on the system configurations.
- Adapter module
Adapter module is handling the complexity of integrating different data formats and storing them in the way that easy to search and retrieve.

\begin{figure}[htp]
    \centering
    \includegraphics[width=1\textwidth]{soa/soa_v1.jpg}
    \caption{Proposed Service Oriented Architecture.}
    \label{fi:proposed_soa}
\end{figure}

>>>>

\section{Service Oriented Architecture}
\dbc{Use a more meaningful title that reflect the solution. Also, why are we talking about an initial phase? You can mention that we iterate through several designs. But it's not important to discuss each design in detail unless once design is essential to understand the other design. Ideally, present only the finalized design while adding a single para for each of the previous attempts.}

In the initial phase of WDIAS architecture, it proposed with a \acrfull{soa} as shown in the \ref{fi:proposed_soa}. Mainly it consist of 4 modules, namely:
- Import modules
Import modules are using for integration of data from different source which come in different data formats.
- Export modules
Export modules are using for data dissemination with different format which may required for different type of operations.
- Extension modules
Extension modules are capable of providing the data assimilation functionalities by providing some of key functionalities such as transformation, interpolation and validation. While data assimilation, existing timeseries may be modified or create new timeseries based on the system configurations.
- Adapter module
Adapter module is handling the complexity of integrating different data formats and storing them in the way that easy to search and retrieve.

\begin{figure}[htp]
    \centering
    \includegraphics[width=1\textwidth]{soa/soa_v1.jpg}
    \caption{Service oriented architecture of WADIS.}
    \label{fi:proposed_soa}
\end{figure}

### 2020-02-09 method/sec_3-microservice.tex
Added Database Structure details
Paper - update abstract

@@

<add the API description and how it related to moduels>
- Import
- Export
- Timeseries
- Extensions

>>>>

- Brief introduction to the Kubernetes

\begin{figure}[htp]
    \centering
    \includegraphics[width=1\textwidth]{method/microservice/microservice_architecture-handle_on_async-v3.jpg}
    \caption{WDIAS architecture to handle request asynchronously.}
    \label{fi:wdias_micro_async}
\end{figure}

\dbc{Dont' use title case of figure captions. Capitalize only the proper nouns. Also, WDIAS can be shorten in figure captions.}

- Explain how k8s support microservice architecture

@@

<add the API description and how it related to modules>
- Import
- Export
- Timeseries
- Extensions

### 2020-02-09 summary/sec_1-summary.tex
Paper - update abstract

First the architecture starts with Service Oriented Architecture using a \acrfull{esb}, and then moved to Actor model based architecture. After comparing the disadvantages of using such architecture with the system requirements, final architecture come up with modern microservice architecture.
Based on the microservice architecture patterns and the nature of the weather metadata, the \acrshort{wdias} came up with the idea of hierarchical database structure in order to provide higher performance while store the weather data. The system uses \acrshort{influxdb} timeseries database for storing Scalar and Vector timeseries data, and using \acrshort{netcdf} for storing Grid timeseries data.
Also the system provide a generic open mechanism to integrate new modules as extension in order to enhance the features of the system. This capability enable to integrate weather data preprocessing flows as extension modules.
The extension API provide easy access to create and modify the extension triggers on the fly without stopping the system or any downtime in order to change the configurations.
The system provides extensive timeseries query endpoints to easy search over the system timeseries metadata with supporting geo based queries.

>>>>

First the architecture starts with Service Oriented Architecture using a \acrfull{esb}, and then moved to Actor model based architecture. After comparing the disadvantages of using such architecture with the system requirements, final architecture come up with modern microservice architecture.
Based on the microservice architecture patterns and the nature of the weather metadata, the \acrshort{wdias} came up with the idea of hierarchical database structure in order to provide higher performance while store the weather data. The system uses \acrshort{influxdb} timeseries database for storing Scalar and Vector timeseries data, and using \acrshort{netCDF} for storing Grid timeseries data.
Also the system provide a generic open mechanism to integrate new modules as extension in order to enhance the features of the system. This capability enable to integrate weather data preprocessing flows as extension modules.
The extension API provide easy access to create and modify the extension triggers on the fly without stopping the system or any downtime in order to change the configurations.
The system provides extensive timeseries query endpoints to easy search over the system timeseries metadata with supporting Geo based queries.
